### 2019-08-08 tex/gitlinks.tex
misc

\newcommand{\codelink}[1]{\href{https://github.com/user/repo/blob/689c6f0d93a04bf4a19694742579915b3bfc57f2/tex/figures/#1.py}{\codeicon}\,\,}
\newcommand{\animlink}[1]{\href{https://github.com/user/repo/blob/689c6f0d93a04bf4a19694742579915b3bfc57f2/tex/figures/#1.gif}{\animicon}\,\,}
\newcommand{\prooflink}[1]{\href{https://github.com/user/repo/blob/689c6f0d93a04bf4a19694742579915b3bfc57f2/tex/proofs/#1.ipynb}{\raisebox{-0.1em}{\prooficon}}}