### 2019-08-08 tex/gitlinks.tex
working on the paper figs

\newcommand{\codelink}[1]{\href{https://github.com/user/repo/blob/60dfb1cf40302ddb3818411b0925c910a2f52625/tex/figures/#1.py}{\codeicon}\,\,}
\newcommand{\animlink}[1]{\href{https://github.com/user/repo/blob/60dfb1cf40302ddb3818411b0925c910a2f52625/tex/figures/#1.gif}{\animicon}\,\,}
\newcommand{\prooflink}[1]{\href{https://github.com/user/repo/blob/60dfb1cf40302ddb3818411b0925c910a2f52625/tex/proofs/#1.ipynb}{\raisebox{-0.1em}{\prooficon}}}