### 2019-08-21 tex/gitlinks.tex
timing figure generation

\newcommand{\codelink}[1]{\href{https://github.com/user/repo/blob/bc985edc746c392492307b3c3f2ebfe9f0fac247/tex/figures/#1.py}{\codeicon}\,\,}
\newcommand{\animlink}[1]{\href{https://github.com/user/repo/blob/bc985edc746c392492307b3c3f2ebfe9f0fac247/tex/figures/#1.gif}{\animicon}\,\,}
\newcommand{\prooflink}[1]{\href{https://github.com/user/repo/blob/bc985edc746c392492307b3c3f2ebfe9f0fac247/tex/proofs/#1.ipynb}{\raisebox{-0.1em}{\prooficon}}}
\newcommand{\cilink}[1]{\href{https://dev.azure.com/user/repo/_build}{#1}}