### 2019-08-21 tex/gitlinks.tex
gen paper only if tag present

\newcommand{\codelink}[1]{\href{https://github.com/user/repo/blob/3bca8febf1e8738da63bacbc3959e6442e52272c/tex/figures/#1.py}{\codeicon}\,\,}
\newcommand{\animlink}[1]{\href{https://github.com/user/repo/blob/3bca8febf1e8738da63bacbc3959e6442e52272c/tex/figures/#1.gif}{\animicon}\,\,}
\newcommand{\prooflink}[1]{\href{https://github.com/user/repo/blob/3bca8febf1e8738da63bacbc3959e6442e52272c/tex/proofs/#1.ipynb}{\raisebox{-0.1em}{\prooficon}}}
\newcommand{\cilink}[1]{\href{https://dev.azure.com/user/repo/_build}{#1}}